Randomized trace estimation for parameter-dependent matrices applied to spectral density approximation

### 2024-10-02 paper/numresults.tex
downsize and adjust parametres to match Krylov-aware

The Krylov-aware stochastic trace estimation \cite{chen-2023-krylovaware-stochastic} also works well for the approximation of the trace of a matrix function $f(\mtx{A}, t)$ which separately depends on a parameter $t$. It also samples two standard Gaussian random matrices $\mtx{\Omega} \in \mathbb{R}^{n \times n_{\mtx{\Omega}}}$ and $\mtx{\Psi} \in \mathbb{R}^{n \times n_{\mtx{\Psi}}}$, of which $\mtx{\Omega}$ is used to extract a low-rank approximation to $f(\mtx{A}, t)$ from a Krylov-subspace of $\mtx{A}$ and $\mtx{\Psi}$ to estimate the trace of the approximation residual.

@@

\begin{figure}[ht]
    \begin{minipage}[c]{.475\linewidth}
        \centering
        \input{plots/krylov_aware_density.pgf}
    \end{minipage}\hfill%
    \begin{minipage}[c]{.475\linewidth}
        \vspace{-35pt}
        \input{tables/krylov_aware_density_KA.tex}
        \newline
        \vspace{15pt}
        \newline
        \input{tables/krylov_aware_density_CN.tex}
    \end{minipage}
    \caption{For the example from \refsec{subsec:hamiltonian} we compare the Chebyshev-Nyström++ method with the Krylov-Aware stochastic trace estimator in multiple configurations (right). To this extent, we compute the $L^1$-error of the methods for various levels of smoothing $\sigma$ (left).}
    \label{fig:krylov-aware-density}
\end{figure}

@@

\begin{figure}[ht]
    \begin{minipage}[c]{.475\linewidth}
        \centering
        \input{plots/krylov_aware_spin.pgf}
    \end{minipage}\hfill%
    \begin{minipage}[c]{.475\linewidth}
        \vspace{-35pt}
        \input{tables/krylov_aware_spin_KA.tex}
        \newline
        \vspace{15pt}
        \newline
        \input{tables/krylov_aware_spin_CN.tex}
    \end{minipage}
    \caption{For the Heisenberg spin chain example, we compare the Chebyshev-Nyström++ method with the Krylov-Aware stochastic trace estimator in the same configurations as in \cite[Table 5.1]{chen-2023-krylovaware-stochastic} (right). To this extent, we compute the $L^1$-error of the methods for various choices of the temperature parameter $\beta^{-1}$ (left).}
    \label{fig:krylov-aware-spin}
\end{figure}

>>>>

The Krylov-aware stochastic trace estimation \cite{chen-2023-krylovaware-stochastic} also works well for the approximation of the trace of a matrix function $f(\mtx{A}, t)$ which separately depends on a parameter $t$. It also samples two standard Gaussian random matrices $\mtx{\Omega} \in \mathbb{R}^{n \times n_{\mtx{\Omega}}}$ and $\mtx{\Psi} \in \mathbb{R}^{n \times n_{\mtx{\Psi}}}$. $\mtx{\Omega}$ is used to extract a low-rank approximation to $f(\mtx{A}, t)$ from a block-Krylov subspace of $\mtx{A}$ generated from $r + q$ iterations of the block-Lanczos method of which the first $r$ are reorthogonalized. $\mtx{\Psi}$ is used to estimate the trace of the approximation residual with $r$ iterations of the Lanczos method.

@@

\begin{figure}[ht]
    \begin{minipage}[c]{.475\linewidth}
        \centering
        \input{plots/krylov_aware_density.pgf}
    \end{minipage}\hfill%
    \begin{minipage}[c]{.475\linewidth}
        \vspace{-35pt}
        \input{tables/krylov_aware_density_KA.tex}
        \newline
        \vspace{15pt}
        \newline
        \input{tables/krylov_aware_density_CN.tex}
    \end{minipage}
    \caption{For the example from \refsec{subsec:hamiltonian} we compare the Chebyshev-Nyström++ (CN++) method with the Krylov-Aware (KA) stochastic trace estimator in multiple configurations (right). To this extent, we compute the $L^1$-error of the methods for various levels of smoothing $\sigma$ (left).}
    \label{fig:krylov-aware-density}
\end{figure}

@@

\begin{figure}[ht]
    \begin{minipage}[c]{.475\linewidth}
        \centering
        \input{plots/krylov_aware_spin.pgf}
    \end{minipage}\hfill%
    \begin{minipage}[c]{.475\linewidth}
        \vspace{-35pt}
        \input{tables/krylov_aware_spin_KA.tex}
        \newline
        \vspace{15pt}
        \newline
        \input{tables/krylov_aware_spin_CN.tex}
    \end{minipage}
    \caption{For the Heisenberg spin chain example, we compare the Chebyshev-Nyström++ (CN++) method with the Krylov-Aware stochastic trace estimator in the same configurations as in \cite[Table 5.1]{chen-2023-krylovaware-stochastic} (right). To this extent, we compute the $L^1$-error of the methods for various choices of the temperature parameter $\beta^{-1}$ (left).}
    \label{fig:krylov-aware-spin}
\end{figure}
